---
title: "Introduction to DTwrappers"
output: rmarkdown::html_vignette
author: 'Authors: David Shilane, Caffrey Lee, Zoe Huang, Anderson Nelson'
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Introduction to DTwrappers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r library, echo=FALSE}
library(formulaic)
library(data.table)
library(knitr)
library(DTwrappers)
data("snack.dat", package = "formulaic")
```

```{r constant,echo=FALSE}
id.name <- "User ID"
awareness.name <- "Awareness"
satisfaction.name <- "Satisfaction"
age.name <- "Age"
gender.name <- "Gender"
income.name <- "Income"
region.name <- "Region"
persona.name <- "Persona"
bp.patterns <- "BP_"
consumption.name <- "Consumption"
consideration.name <- "Consideration"
advocacy.name <- "Advocacy"
satisfaction.name <- "Satisfaction"
age.group.name <- "Age Group"
income.group.name <- "Income Group"
```

# Introduction

Many newcomers to the data.table package are overwhelmed by the unique syntax . **DTwrappers** is useful in applying data.table with wrapper functions while still maintaining the data.table execution speed. 

**DTwrappers** package has 9 main functions. The main purpose of developing the package is to help users to apply data.table package faster and more convenient.

 - **create.filter.expression** automatically creates filter expressions from a provided a character value, logical value, or expression stating the logical operations.

```{r create.filter.expression_example1}
create.filter.expression(the.filter = NULL)
```

```{r create.filter.expression_example2}
create.filter.expression(the.filter = c(age.name, region.name))
```

```{r create.filter.expression_example3}
create.filter.expression(the.filter = age.name == region.name)
```

```{r create.filter.expression_example4}
create.filter.expression(the.filter = "get(region.name) == 'South'")
```

```{r create.filter.expression_example5}
create.filter.expression(the.filter = "get(region.name) == 'South' & get(age.name) > 20")
```

```{r create.filter.expression_example1}
create.filter.expression(the.filter = "get(region.name) == 'South' & get(age.name) > 20")
```

 - **dt.count** is a wrapper to computations like `dat[the.filter, .N, keyby = grouping.variables]` that counts the number of rows that meets the qualification such as filter and the by (grouping variables).

```{r dt.count_example1}
dt.count(dat = snack.dat)
```

```{r dt.count_example2}
dt.count(dat = snack.dat,
grouping.variables = c(region.name, gender.name))
```

```{r dt.count_example3}
dt.count(dat = snack.dat,
the.filter = "Age > 65",
grouping.variables = c(region.name, gender.name),
count.name = "Records with Age > 65")
```

```{r dt.count_example4}
dt.count(dat = snack.dat,
the.filter = 'get(age.name) > 20',
grouping.variables = c(age.name,income.name),
grouping.type = 'keyby', count.name = "counts")
```
 
 - **dt.filter** subsets the data corresponding to the filtering statement.
 
 - **dt.select** provides more flexibility in selecting the rows to include such as first.k, last.k, or specific row by setting up the parameters.
 
 - **df.define.variable** is a wrapper to computations like `dat[, new.variable := new.values]` that allows users to re-define the variables so that users have freedom to set the variable name and the the values. The values could be either expression, direct code(data.table format), or character.

```{r df.define.variable_example1}
snack.dat1 = copy(snack.dat)
snack.dat1 <- dt.define.variable(dat = snack.dat1,
variable.name = "Age Decade",
the.values = snack.dat1[, floor(get(age.name) / 10)])
snack.dat1[1:10, .SD, .SDcols = c(age.name, "Age Decade")]
```

```{r df.define.variable_example2}
snack.dat2 = copy(snack.dat)
snack.dat2 <- dt.define.variable(dat = snack.dat2,
variable.name = "Income in Thousands",
the.values = expression(floor(get(income.name) / 10^3)),
specification = "by.expression")
snack.dat2[1:10, .SD, .SDcols = c(income.name, "Income in Thousands")]
```

```{r df.define.variable_example3}
snack.dat3 = copy(snack.dat)
snack.dat3 <- dt.define.variable(dat = snack.dat3,
variable.name = "Income in Thousands",
the.values = "floor(get(income.name) / 10^3)",
specification = "by.expression")
snack.dat3[1:10, .SD, .SDcols = c(income.name, "Income in Thousands")]
```

```{r df.define.variable_example4}
snack.dat4 = copy(snack.dat)
snack.dat4 <- dt.define.variable(dat = snack.dat4,
variable.name = "Region and Country",
the.values = expression(sprintf('%s, USA', get(region.name))),
specification = "by.expression")
snack.dat4[1:10, .SD, .SDcols = c(region.name, "Region and Country")]
```
 
 - **dt.first.k.rows** returns the first K number of rows from the given data frame based on the variables and the grouping varaibles statement.
 
 - **dt.last.k.rows** returns the last K number of rows from the given data frame based on the variables and the grouping varaibles statement.
 
 - **dt.lapply** is a wrapper to calculations like `dat[the.filter, lapply(X = .SD, FUN = function.name), .SDcols = the.variables, keyby = grouping.variables]` by specifying:
 
    - the.filter (character or expression) for the i step
    
    - the.variables (character vector of column names) for the .SDcols
    
    - the.functions (character vector of function names) for the functions to apply
    
    - grouping.variables (character vector of column names) for the variables to group by.
 
 - **dt.remove.variables** is a wrapper to computations like `dat[, this.variable := NULL]` that removes selected columns from dat
 
 